---
title: "Introduction to saphy"
author: "The Hackout3 Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A time-stamped phylogeny of Zika virus is included in the package directory.

```{r,echo=FALSE}
suppressPackageStartupMessages(library(ape))
suppressPackageStartupMessages(library(saphy))
suppressPackageStartupMessages(library(adephylo))
```

```{r}
tr <- read.tree(file=paste(path.package("saphy"),"/trees/Zika_newick.tre",sep=""))
```

```{r}
plot(tr)
```


As this is a rooted, time-stamped phylogeny, we can get (relative) sampling times by calculating the distance of each tip to the root, using the ```distRoot``` function from the ```adephylo``` package. This is done automatically, when we prune the tree time-wise using the ```timeprune``` function, without specifying sampling times.

```{r}
tr.pruned <- timeprune(tr)
```

This returns a list with two elements; ```tipdates```, a named vector of tip dates/times, and ```trees```, a list of ```phylo``` objects (named by the newest tip).

```{r}
names(tr.pruned)
```

The tip length associated with each new sequence can be calculated using the ```tiplength``` function.

```{r}
tl <- mapply(tiplength,tr.pruned$trees,names(tr.pruned$trees))
```

We can now plot the tip lengths over sampling time; we drop the sampling time of the first sequence.

```{r}
plot(tl~tr.pruned$tipdates[-1],type="h",xlab="Distance to root",ylab="Tip length")
```


